Vertical shear instability in astrophysical disks: vertically global linear calculations with thermal relaxation and applications to the Solar Nebula

### 2015-02-27 linear_iso.tex
minor changes
updated all figs

We further discuss caveats of the VGSB in Appendix \ref{global_corr}.
However, we find that for perturbations with small radial lengthscales
(large $|k_x|$) and fast thermal relaxation ($\beta\ll1$), which
characterizes the VSI, the VGSB framework is adequate for the purpose
of a simple analytical discussion.

@@

\subsection{Isothermal perturbations}\label{iso_discuss}
Isothermal perturbations correspond to $\beta\equiv0$, 
%relaxation timescale $\be
%or for any adiabatc index $\gamma$ when
%$\beta=0$. 
so $\chi=1$ and Eq. \ref{ode_Q} implies 
\begin{align}
  Q = W, 
\end{align}
 and so 
 \begin{align}
   \ii\hat{\sigma}c_s\delta v_z = W^\prime
   % \quad   \text{(isothermal perturbations)},
 \end{align}
from Eq. \ref{lin_vz}. For isothermal perturbations 
the parameter $\chi=1$ so it is easiest to work with
$W$ by integrating Eq. \ref{vertiso_gov_nondim}. For later discussion
it is convenient to write the result as

>>>>

We further discuss caveats of the VGSB in Appendix \ref{global_corr}.
However, we find that for perturbations with small radial lengthscales
($k_xH_\mathrm{iso}\gg 1$) and fast thermal relaxation ($\beta\ll1$), which
characterizes the VSI, the VGSB framework is adequate for the purpose
of a simple analytical discussion.

@@

\subsection{Isothermal perturbations}\label{iso_discuss}
Isothermal perturbations correspond to $\beta\equiv0$, 
%relaxation timescale $\be
%or for any adiabatc index $\gamma$ when
%$\beta=0$. 
so $\chi=1$ and Eq. \ref{ode_Q} implies 
\begin{align}
  Q = W, 
\end{align}
 and so 
 \begin{align}
   \ii\hat{\sigma}c_s\delta v_z = W^\prime
   % \quad   \text{(isothermal perturbations)},
 \end{align}
from Eq. \ref{lin_vz}. For isothermal perturbations it is in fact 
simpler to work with $W$ by integrating Eq. \ref{vertiso_gov_nondim}. For later discussion
it is convenient to write the result as

### 2015-02-27 setup.tex
updated all figs

\subsection{Radially-local approximation}  
We consider perturbations with small radial lengthscales 
so that we can ignore the radial variations in the background
structure and radial gradients. We thus evaluate all coefficients in
the above equations at the fiducial radius $r=r_0$. However, the full
vertical dependence of the coefficients are retained. Then we can
assume perturbations have spatial dependence in the form
\begin{align}
  \delta \rho(r,z) \to \delta\rho(z)\exp{\ii k_r r},
\end{align}
and we assume $|k_rr|\gg1$ so that curvature terms can also be
ignored.

In order to emphasize the radially-local approximation, we
introduce the local Cartesian co-ordinate $(x,y,z)$ corresponding to
the the local $(r,\phi,z)$ directions, respectively. Then writing
$\delta P /\rho \equiv W$ and $c_s^2\delta\rho/\rho\equiv Q$, the
linearized system of equations are 
\begin{align}
  \ii\sigma \frac{Q}{c_s^2}  &=  \ii k_x \delta v_x + \frac{d\dd
    v_z}{dz} + \frac{\p\ln{\rho}}{\p z}\delta v_z + \hat{g}_c
  \frac{\p\ln{\rho}}{\p r} \delta v_x,\label{lin_mass}\\
  \ii\sigma \delta v_x  &= \ii k_x W - 2\Omega\delta v_y -
  \hat{g}_c\frac{1}{\rho}\frac{\p P}{\p r}\frac{Q}{c_s^2}\label{lin_vx}\\
   \ii \sigma\delta v_y &= r\frac{\p \Omega}{\p z}\delta v_z +
  \frac{\kappa^2}{2\Omega}\delta v_x, \label{lin_vy}\\
   \ii\sigma\delta v_z &= \frac{dW}{dz} +
  \frac{\p\ln{\rho}}{\p z}\left(W-Q\right), \label{lin_vz}\\
  \ii \sigma W &= c_s^2\frac{\p\ln{\rho}}{\p z}\delta v_z +
  c_s^2\frac{\gamma}{\Gamma} \left(\ii k_x\delta v_x + \frac{d\delta
      v_z}{dz}\right) \notag\\
  &+ \frac{1}{t_c}\left(W - \frac{Q}{\Gamma}\right) +
  \hat{g}_c\frac{1}{\rho}\frac{\p P}{\p r}\delta v_x.\label{lin_energy}
  % -&\ii\sigma W + \frac{1}{t_c}\left(W-\frac{Q}{\Gamma}\right) = -\ii\sigma\frac{\gamma}{\Gamma} Q +
  % c_s^2\frac{d\ln{\rho}}{dz}\left(\frac{\gamma}{\Gamma}-1\right)\delta v_z.\label{lin_energy}
\end{align}
It is understood that the quantities $\rho,\,c_s^2,\,\Omega,\kappa,\, 
\p_z\Omega,\, \p_r\ln{\rho} $ and $ \p_r P$ are evaluated at $r=r_0$.   
Eq. \ref{lin_mass}---\ref{lin_energy} will form the basis of our numerical
study, but for analytical discussion
we will make further simplifications.

>>>>

\subsection{Radially-local approximation}  
We consider perturbations with small radial lengthscales 
so that we can ignore the radial variations in the background
structure and radial gradients. We thus evaluate all coefficients in
the above equations at the fiducial radius $r=r_0$. However, the full
vertical dependence of the coefficients are retained. Then we can
assume perturbations have spatial dependence in the form
\begin{align}
  \delta \rho(r,z) \to \delta\rho(z)\exp{\ii k_r r}.
\end{align}
We take $k_r>0$ without loss of generality, and further assume $k_rr\gg1$ so that curvature terms can also be 
ignored.

In order to emphasize the radially-local approximation, we
introduce the local Cartesian co-ordinate $(x,y,z)$ corresponding to
the the local $(r,\phi,z)$ directions, respectively. Then writing
$\delta P /\rho \equiv W$ and $c_s^2\delta\rho/\rho\equiv Q$, the
linearized system of equations are 
\begin{align}
  \ii\sigma \frac{Q}{c_s^2}  &=  \ii k_x \delta v_x + \frac{d\dd
    v_z}{dz} + \frac{\p\ln{\rho}}{\p z}\delta v_z + \hat{g}_c
  \frac{\p\ln{\rho}}{\p r} \delta v_x,\label{lin_mass}\\
  \ii\sigma \delta v_x  &= \ii k_x W - 2\Omega\delta v_y -
  \hat{g}_c\frac{1}{\rho}\frac{\p P}{\p r}\frac{Q}{c_s^2}\label{lin_vx}\\
   \ii \sigma\delta v_y &= r\frac{\p \Omega}{\p z}\delta v_z +
  \frac{\kappa^2}{2\Omega}\delta v_x, \label{lin_vy}\\
   \ii\sigma\delta v_z &= \frac{dW}{dz} +
  \frac{\p\ln{\rho}}{\p z}\left(W-Q\right), \label{lin_vz}\\
  \ii \sigma W &= c_s^2\frac{\p\ln{\rho}}{\p z}\delta v_z +
  c_s^2\frac{\gamma}{\Gamma} \left(\ii k_x\delta v_x + \frac{d\delta
      v_z}{dz}\right) \notag\\
  &+ \frac{1}{t_c}\left(W - \frac{Q}{\Gamma}\right) +
  \hat{g}_c\frac{1}{\rho}\frac{\p P}{\p r}\delta v_x,\label{lin_energy}
  % -&\ii\sigma W + \frac{1}{t_c}\left(W-\frac{Q}{\Gamma}\right) = -\ii\sigma\frac{\gamma}{\Gamma} Q +
  % c_s^2\frac{d\ln{\rho}}{dz}\left(\frac{\gamma}{\Gamma}-1\right)\delta v_z.\label{lin_energy}
\end{align}
where we have made the notation change $k_r\to k_x$. 
It is understood that the quantities $\rho,\,c_s^2,\,\Omega,\kappa,\, 
\p_z\Omega,\, \p_r\ln{\rho} $ and $ \p_r P$ are evaluated at $r=r_0$ using their expressions in the global disk. 
Eq. \ref{lin_mass}---\ref{lin_energy} will form the basis of our numerical
study, but for analytical discussion
we will make further simplifications.
